---
title: "Data Analysis Assignment"
author: "Pardeepan"
date: "2023-09-10"
output: html_document
---

# Introduction

This is my data analysis assignment, where I will analyze the "Car_sales.csv" dataset using R.

## Dataset

I have loaded the dataset into a variable named 'df' for a data frame.

```{r}
# Load the dataset into a variable (e.g., df for a data frame)
df <- read.csv("Car_sales.csv")

# Descriptive statistics for quantitative variables
summary(df$Engine_size)
summary(df$Horsepower)
summary(df$Fuel_efficiency)
summary(df$Price_in_thousands)

# Descriptive statistics for categorical variables
table(df$Manufacturer)
table(df$Vehicle_type)
table(df$Engine_size)

# Logarithmic transformation
df$log_Price_in_thousands <- log(df$Price_in_thousands)

# Descriptive statistics for the transformed variable
summary(df$log_Price_in_thousands)

```{r}
# Load the ggplot2 library if not already loaded
library(ggplot2)

# Histogram for 'Price_in_thousands' variable
ggplot(df, aes(x = Price_in_thousands)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  labs(title = "Distribution of Price_in_thousands", x = "Price_in_thousands", y = "Frequency")

# Scatterplot between 'Price_in_thousands' and 'Model'
ggplot(df, aes(x = Price_in_thousands, y = Model)) +
  geom_point() +
  labs(title = "Scatterplot: Price_in_thousands vs. Model", x = "Price_in_thousands", y = "Model")
